---
title: "CSLS_chem_lake_gw"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{CSLS_chem_lake_gw}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(CSLSfluxes)
library(dplyr)
library(stringr)
library(lubridate)
library(reshape2)
library(ggplot2)
library(extrafont)

text_size   <- 12
lakes       <- c("Pleasant", "Long", "Plainfield")
water_chem  <- CSLSdata::water_chem
water_chem  <- water_chem %>%
               filter(!site_type %in% c("deep"))
water_chem$lake <- factor(water_chem$lake, levels = c("Precip", lakes))
```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_summary <- function(df,
                        yaxis_type = "number",
                        text_size = 12,
                        site_types = c("precipitation", "upgradient", "nogradient", 
                                       "downgradient", "lake", "lake_bottom"),
                        site_labels = c("Precip", "Upgr.", "Nogr.", 
                                        "Downgr.", "Lake Surf", "Lake Bot."),
                        site_colors = c("#1F78B4", "#33A02C", "#B15928", 
                                        "#B2DF8A", "#FB9A99", "#E31A1C")){

  df$site_type <- factor(df$site_type, levels = site_types)
  ylabel       <- sprintf("%s (%s)",
                          unique(df$name),
                          str_to_lower(unique(df$units)))

  plot_obj <- ggplot(df,
                     aes(x = .data$lake,
                         y = .data$result,
                         group = .data$site_type,
                         color = .data$site_type,
                         fill = .data$site_type)) +
              geom_hline(aes(yintercept = min(.data$loq, na.rm = TRUE),
                             linetype = "LOQ"),
                         color = "grey40",
                         size = 1) +
              geom_hline(aes(yintercept = min(.data$lod, na.rm = TRUE),
                             linetype = "LOD"),
                         color = "grey40",
                         size = 1) +
              geom_point(position = position_jitterdodge(jitter.width = 0.1,
                                                         dodge.width = 0.4),
                         size = 2.5,
                         color = "grey70") +
              stat_summary(fun.y = mean,
                           fun.ymin = min,
                           fun.ymax = max,
                           shape = "\U25AC",
                           size = 1,
                           position = position_dodge(width = 0.4)) +
              labs(x = "", y = ylabel, title = "Everything", 
                   color = "Site", fill = "Site") +
              scale_linetype_manual(name = "",
                                    breaks = c("LOQ", "LOD"),
                                    values = c("longdash", "dotted")) +
              scale_color_manual(name = "Sites",
                                 breaks = site_types,
                                 limits = site_types,
                                 labels = site_labels,
                                 values = site_colors) +
              guides(color = guide_legend(order = 1),
                     fill = FALSE,
                     linetype = guide_legend(order = 2)) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    panel.grid.major = element_blank(),
                    panel.grid.minor = element_blank(),
                    legend.position = "top",
                    plot.title = element_text(hjust = 0.5))

  if (yaxis_type == "log") {
    plot_obj <- plot_obj + scale_y_log10()
  }

  return(plot_obj)
}


```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_timeseries <- function(df,
                            yaxis_type = "number",
                            text_size = 12,
                            site_types = c("upgradient",
                                           "nogradient",
                                           "downgradient"),
                            site_labels = c("Upgr.", "Nogr.", "Downgr."),
                            site_colors = c("#33A02C", "#B15928", "#B2DF8A"),
                            site_shapes = c(21, 21, 21)){
  ylabel   <- sprintf("%s (%s)",
                      unique(df$name),
                      str_to_lower(unique(df$units)))
  df       <- df %>% arrange(.data$date)

  plot_obj <- ggplot(df,
                     aes(x = floor_date(.data$date, unit = "day"),
                         y = .data$result,
                         group = .data$site_id,
                         color = .data$site_type,
                         fill = .data$site_type,
                         shape = .data$site_type)) +
              geom_hline(aes(yintercept = min(.data$loq, na.rm = TRUE),
                             linetype = "LOQ"),
                         color = "grey40",
                         size = 1) +
              geom_hline(aes(yintercept = min(.data$lod, na.rm = TRUE),
                             linetype = "LOD"),
                         color = "grey40",
                         size = 1) +
              geom_path() +
              geom_point(size = 2.5) +
              facet_grid(~lake) +
              labs(x = "", y = ylabel, title = "Groundwater", color = "Site") +
              scale_x_datetime(breaks = "6 months",
                               date_labels = "%b %y") +
              scale_linetype_manual(name = "",
                                    breaks = c("LOQ", "LOD"),
                                    values = c("longdash", "dotted")) +
              scale_color_manual(name = "Sites",
                                 breaks = site_types,
                                 limits = site_types,
                                 labels = site_labels,
                                 values = site_colors) +
              scale_fill_manual(name = "Sites",
                                breaks = site_types,
                                limits = site_types,
                                labels = site_labels,
                                values = site_colors) +
              scale_shape_manual(name = "Sites",
                                 breaks = site_types,
                                 limits = site_types,
                                 labels = site_labels,
                                 values = site_shapes) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    legend.position = "top",
                    plot.title = element_text(hjust = 0.5))
  if (yaxis_type == "log") {
    plot_obj <- plot_obj + scale_y_log10()
  }
  return(plot_obj)
}

```

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
plot_timeseries_by_year <- function(df,
                            yaxis_type = "number",
                            text_size = 12){
  ylabel   <- sprintf("%s (%s)",
                      unique(df$name),
                      str_to_lower(unique(df$units)))
  df$year  <- as.factor(year(df$date))
  df$date  <- yday(df$date)
  df       <- df %>% arrange(.data$date)
  path_df  <- df %>%
              filter(.data$site_type %in% c("lake", "upgradient", "downgradient"))

  plot_obj <- ggplot(df,
                     aes(x = as.Date(.data$date, origin = as.Date("2018-01-01")),
                         y = .data$result,
                         group = .data$year,
                         color = .data$year,
                         fill = .data$year,
                         shape = .data$year)) +
              geom_hline(aes(yintercept = min(.data$loq, na.rm = TRUE),
                             linetype = "LOQ"),
                         color = "grey40",
                         size = 1) +
              geom_hline(aes(yintercept = min(.data$lod, na.rm = TRUE),
                             linetype = "LOD"),
                         color = "grey40",
                         size = 1) +
              geom_path(data = path_df) +
              geom_point(size = 2.5) +
              facet_grid(~lake) +
              labs(x = "", y = ylabel, title = "Lake Surface", color = "Site") +
              scale_x_date(date_breaks = "2 months",
                           date_labels = "%b") +
              scale_linetype_manual(name = "",
                                    breaks = c("LOQ", "LOD"),
                                    values = c("longdash", "dotted")) +
              scale_color_brewer(name = "Year",
                                 palette = "Paired") +
              scale_fill_brewer(name = "Year",
                                palette = "Paired") +
              scale_shape_discrete(name = "Year") +
              guides(linetype = guide_legend(order = 2), 
                     color = guide_legend(order = 1), 
                     fill = guide_legend(order = 1), 
                     shape = guide_legend(order = 1)) +
              theme_bw() +
              theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size),
                    legend.position = "top",
                    plot.title = element_text(hjust = 0.5))
  if (yaxis_type == "log") {
    plot_obj <- plot_obj + scale_y_log10()
  }
  return(plot_obj)
}

```

# Overview {#top}
The data used here includes data from both SWIMS and LDES, with most
precipitation data drawn from NADP. Duplicates between SWIMS and LDES as well as
samples labeled as blanks or duplicates are removed.

<br>

# Acidity

## pH
[Back to top](#top)

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"PH LAB"
plotting_name <- "pH"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Alkalinity, Total CaCO3
[Back to top](#top)


```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <- c("ALKALINITY TOTAL CACO3", "ALKALINITY TOTAL GRAN AVAL UEQ/L")
plotting_name <- "Alkalinity Total CaCO3"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Salinity

## Conductivity
[Back to top](#top)

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"CONDUCTIVITY, UMHOS/CM @ 25C"
plotting_name <- "Conductivity"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Major Cations

## Calcium, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high Ca
(>> 100 mg/L).
  
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"

this_parameter <- filter_parameter(water_chem, 
                                   parameter, 
                                   plotting_name,
                                   no_bad_well = TRUE,
                                   note_lake_bottom = TRUE)
this_gw    <- this_parameter %>% 
              filter(.data$site_type %in% 
                       c("upgradient", "nogradient", "downgradient"))
this_lake <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Magnesium, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high Mg
(>> 100 mg/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   no_bad_well = TRUE,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Sodium, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PSNT-06, which has high Na
(> 60 mg/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"SODIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Sodium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-06")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake") 

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Potassium, Total Recoverable
[Back to top](#top)


```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"POTASSIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Potassium"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Ca:Mg Ratio
[Back to top](#top)

Notes:

* Included here: GW Monitoring Well PFL-09, which has very high Ca & Mg (>> 100
mg/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"CALCIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Calcium"
Ca            <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
Ca            <- Ca %>%
                 select(.data$lake, .data$site_type, .data$date, .data$result,
                        .data$site_id)

parameter     <- "MAGNESIUM TOTAL RECOVERABLE"
plotting_name <- "Total Recoverabel Magnesium"
Mg            <- filter_parameter(water_chem, parameter, plotting_name)
Mg            <- Mg %>%
                 select(.data$lake, .data$site_type, .data$date, .data$result,
                        .data$site_id)

Ca_Mg         <- merge(Ca, Mg, by = c("date", "lake", "site_type", "site_id"))
Ca_Mg$result  <- Ca_Mg$result.x/Ca_Mg$result.y
Ca_Mg$lod     <- 0
Ca_Mg$loq     <- 0

this_gw        <- Ca_Mg %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- Ca_Mg %>% filter(.data$site_type == "lake")

plot_summary(Ca_Mg, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Hardness, Total Recoverable Calculation
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high
hardness (>> 1000 mg/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"HARDNESS TOTAL RECOVERABLE CALCULATION"
plotting_name <- "Total Recoverable Hardness"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   no_bad_well = TRUE,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Major Anions

## Chloride
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PSNT-06, which has high Cl
(30-170 mg/L).


```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"CHLORIDE"
plotting_name <- "Chloride"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-06")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Sulfate, Total
[Back to top](#top)

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <- c("SULFATE TOTAL", "SULFATE DISS, AS SO4")
plotting_name <- "Sulfate"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Nutrients

## Phosphorus, Total
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which gets to very high
TP (10 mg/L) and GW Monitoring Well PSNT-05, which also has high TP (0.25-2.5 mg/L)
* Otherwise, groundwater levels are typically similar to or slightly higher than
lake levels
* Lake values are often below the limit of quality, particularly at Pleasant
Lake.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"PHOSPHORUS TOTAL"
plotting_name <- "Total Phosphorus"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   no_bad_well = TRUE,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-05")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake") 
this_parameter <- filter_parameter(water_chem, parameter, plotting_name)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Nitrogen, NH3-N Dissolved
[Back to top](#top)

Notes:

* NH3-N is behaves similarly at all three lakes
* NH3-N is almost non-existant in groundwater (at/below the limit of quality in
all but two samples) and still very low in lakes

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"NITROGEN NH3-N DISS"
plotting_name <- "Dissolved Ammonium as Nitrogen"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Nitrogen, NO3 + NO2 Dissolved (as N)
[Back to top](#top)

Notes:

* All sites have at least one well with appreciable NO3/NO2, but most
groundwater and lake samples are near the limit of quality.
* Pleasant Lake has slightly higher (i.e., usually measureable) lake values
than Long and Plainfield, which are all non-detect.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"NITROGEN NO3+NO2 DISS (AS N)"
plotting_name <- "Dissolved Nitrate/Nitrite as Nitrogen"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Metals

## Aluminum, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high Al
(>> 20,000 ug/L) and PSNT-05, which also has high Al (2,500-20,000 ug/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"ALUMINUM,TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Aluminum"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   no_bad_well = TRUE,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-05")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Iron, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high Fe
(20-400 mg/L) and PSNT-05, which also has high Fe (10-50 mg/L).
* Groundwater levels are generally low (near or below the limit of quality) with
the exception of one well at Pleasant (PSNT-05) and one well at Plainfield
(PFL-09).
* Iron was only measured at the lake surfaces one time (in Feb 2019) and was
below detection at Pleasant, below the limit of quality at Plainfield, and less
than 1 mg/L at Long.

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"IRON TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Iron"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE,
                                   no_bad_well = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-05")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

## Manganese, Total Recoverable
[Back to top](#top)

Notes:

* Not shown in time series: GW Monitoring Well PFL-09, which has very high Mn
(1,000-10,000 ug/L) and PSNT-05, which also has high Mn (up to 3,000 ug/L).

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"MANGANESE, TOTAL RECOVERABLE"
plotting_name <- "Total Recoverable Manganese"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE,
                                   no_bad_well = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient")) %>%
                  filter(.data$site_id != "PSNT-05")
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")
this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```

<br>

# Other

## Silica, Dissolved (as SiO2)
[Back to top](#top)

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4, fig.align="center"}
parameter     <-"SILICA, DISSOLVED (MG/L AS SI02)"
plotting_name <- "Dissolved Silica as SiO2"

this_parameter <- filter_parameter(water_chem, parameter, plotting_name,
                                   note_lake_bottom = TRUE)
this_gw        <- this_parameter %>% 
                  filter(.data$site_type %in% 
                           c("upgradient", "nogradient", "downgradient"))
this_lake      <- this_parameter %>% filter(.data$site_type == "lake")

plot_summary(this_parameter, yaxis_type = "log", text_size)
plot_timeseries(this_gw, yaxis_type = "number", text_size)
plot_timeseries_by_year(this_lake, text_size = text_size)
```


