---
title: "MODFLOW_chem_eval"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{MODFLOW_chem_eval}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
library(CSLSfluxes)
library(reshape2)
library(ggplot2)
library(NISTunits)
library(extrafont)
library(lubridate)
library(DT)

lakes     <- c("Pleasant", "Long", "Plainfield")
scenarios <- c("no_irr", "cur_irr")
text_size <- 12

# Exceedance Probabilities
exceeds     <- CSLSscenarios::MODFLOW_metrics %>%
               filter(.data$sim == 1,
                      .data$scenario %in% scenarios,
                      .data$metric == "exceedance_level",
                      .data$series == "month") %>%
               select(lake = .data$lake,
                      probs = .data$variable,
                      yintercept = .data$value,
                      scenario = .data$scenario)
bold_line   <- exceeds %>% filter(.data$probs %in% c("50"))
solid_line  <- exceeds %>% filter(.data$probs %in% c("10","90"))
dashed_line <- exceeds %>% filter(.data$probs %in% c("25","75"))
```


# MODFLOW Water Balance Dynamics

For reference, here are the annual groundwater inflow (% of P+GWin), groundwater
outflow (% of P+GWin), and lake water residence times (Vlake/(P+GWin)) under
recent climate conditions, with "no irrigation" and "current irrigation" land
use conditions.


Note that this is NOT a recreation of history - years refer to climate
conditions only, not actual or projected lake levels in the past.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3, fig.width=6.5}
MODFLOW <- CSLSdata::MODFLOW %>%
           filter(!is.na(.data$GWin_m3),
                  .data$scenario  %in% scenarios,
                  .data$sim == 1,
                  year(.data$date) >= 1986) %>%
           mutate(GWin_m3 = -GWin_m3) %>%
           select(.data$lake, .data$date, .data$level_m, .data$P_m3, .data$E_m3,
                  .data$GWin_m3, .data$GWout_m3, .data$dV_m3, .data$scenario)

add_vol <- NULL
for (lake in lakes) {
  this_lake        <- filter(MODFLOW, .data$lake == !!lake)
  this_bathy       <- filter(CSLSdata::bathymetry, .data$lake == !!lake)
  fn_elev_vol      <- approxfun(this_bathy$elev_m, this_bathy$vol_m3)
  this_lake$vol_m3 <- fn_elev_vol(this_lake$level_m)
  add_vol <- rbind(add_vol, this_lake)
}

MODFLOW <- add_vol %>%
           mutate(GWout_pcnt = 100*.data$GWout_m3/(.data$P_m3 + .data$GWin_m3),
                  GWin_pcnt = 100*.data$GWin_m3/(.data$P_m3 + .data$GWin_m3),
                  res_time_yr = .data$vol_m3/(12*(.data$P_m3 + .data$GWin_m3)))

MODFLOW_yr <- MODFLOW %>%
              group_by(lake = .data$lake,
                       scenario = .data$scenario,
                       year = year(.data$date)) %>%
              summarise(GWout_m3 = sum(.data$GWout_m3),
                        GWin_m3 = sum(.data$GWin_m3),
                        P_m3 = sum(.data$P_m3),
                        vol_m3 = mean(.data$vol_m3, na.rm = TRUE),
                        .groups = "drop") %>% 
              mutate(GWout_pcnt = 100*.data$GWout_m3/(.data$P_m3 + .data$GWin_m3),
                     GWin_pcnt = 100*.data$GWin_m3/(.data$P_m3 + .data$GWin_m3),
                     res_time_yr = .data$vol_m3/(.data$P_m3 + .data$GWin_m3))

ggplot(MODFLOW_yr) +
  geom_line(aes(x = .data$year-1986,
                y = .data$GWin_pcnt,
                group = .data$scenario,
                color = .data$scenario)) +
  facet_grid(~lake) +
  scale_color_manual(name = "",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  labs(x = "", y = "Groundwater Inflow (%)", linetype = "") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")

ggplot(MODFLOW_yr) +
  geom_line(aes(x = .data$year-1986,
                y = .data$GWout_pcnt,
                 group = .data$scenario,
                 color = .data$scenario)) +
  scale_color_manual(name = "",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  facet_grid(~lake) +
  scale_y_continuous(limits = c(0,100),
                     breaks = seq(0,100,20),
                     expand = c(0,0)) +
  labs(x = "", y = "Groundwater Outflow (%)", linetype = "") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")

ggplot(MODFLOW_yr) +
  geom_line(aes(x = .data$year-1986,
                y = .data$res_time_yr,
                 group = .data$scenario,
                 color = .data$scenario)) +
  facet_grid(~lake) +
  scale_color_manual(name = "",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  labs(x = "", y = "Lake Residence Time (yr)", linetype = "") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")
```

<br>

# MODFLOW Mg Budget

## 1. Tune groundwater inflow concentration

### Theory

We start with Magnesium, which should be very conservative, so the following
(very simple) mass-balance equations should hold:

$$ \Delta Storage = In - Out $$
$$ M_{lake2} - M_{lake1} = M_{GWin} + M_{Precip} - M_{GWout}$$
$$ (C_{lake2}*V_{lake2} ) = (C_{lake1}*V_{lake1} )+ (C_{GWin}*V_{GWin}) + (C_{Precip}*V_{Precip}) - (C_{lake1}*V_{GWout})$$

$$C_{GWin} = \frac{(C_{lake2}*V_{lake2}) - (C_{lake1}*V_{lake1}) -(C_{Precip}*V_{Precip}) + (C_{GWout}*V_{GWout})}{V_{GWin}}$$

<br>

### Calculated Groundwater Inflow Concentrations

We make the following assumptions about volumes, fluxes, and concentrations:

* **Volumes and fluxes**. We use values from the MODFLOW calibration simulation
for WY2018 (10/2017 through 09/2018).
* **Precipitation concentration.** We use the median concentration across all
measurements (2014-2019).
* **Starting and ending lake concentrations.** We used the median value in
Sept/Oct 2018 for the starting lake concentration and the median value in
Sept/Oct 2019 as the ending lake concentration.
* **Groundwater outflow concentration.** We use the median lake concentration
across all measurements (05/2018-11/2019).

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
param   <- "MAGNESIUM TOTAL RECOVERABLE"
lakes   <- c("Pleasant", "Long", "Plainfield")

C_pcpn  <- get_C_pcpn(param, lakes = lakes)
C_GWout <- get_C_lake(param,
                      summary_type = "median",
                      value_name = "C_GWout")

C1_lake <- get_C_lake(param, 
                      start_date = as_datetime("2018-09-01"),
                      end_date = as_datetime("2018-10-31"),
                      summary_type = "median",
                      value_name = "C1_lake")
C2_lake <- get_C_lake(param, 
                      start_date = as_datetime("2019-09-01"),
                      end_date = as_datetime("2019-10-31"),
                      summary_type = "median",
                      value_name = "C2_lake")
C_vals  <- left_join(C_pcpn, C1_lake, by = "lake") %>%
           left_join(C2_lake, by = "lake") %>%
           left_join(C_GWout, by = "lake")
C_GWin  <- calculate_C_GWin(C_vals)

datatable(C_GWin)
```

<br>

## 2. Evaluate lake concentrations with and without pumping

Now, we solve for C_lake during the "no irrigation" and "current irrigation"
scenarios, making the following assumptions:

* **Initial lake concentration:** median across all measurements (same as
C_GWout in table above)
* **Precipitation concentration:** median across all measurements (same as C_P
in table above)
* **Groundwater inflow concentration:** as calculated in previous step (same as
C_GWin in table above).

We calculate lake concentration on a daily time step, with monthly MODFLOW
fluxes (P, GWin, GWout) converted to daily values, and lake volumes linearly
interpolated across months. Using a daily time step (as opposed to a monthly or
annual time step) is particularly helpful at Long Lake, where instabilities
(e.g. negative lake concentrations) can result if a time step is too large
compared to the (often very short) lake water residence time.

<br>

### Lake Mg time series
**Important Note** the y-axis is limited to Mg = 40 mg/L, but values at Long Lake can peak to ~165 mg/L.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
C_vals <- C_GWin %>%
          select(lake = .data$lake,
                 C_pcpn = .data$C_pcpn,
                 C0_lake = .data$C_GWout,
                 C_GWin = .data$C_GWin)
C_lake1 <- calculate_C_lake(C_vals,
                            scenario = "no_irr",
                            sim = 1,
                            dt = "day")
C_lake2 <- calculate_C_lake(C_vals,
                            scenario = "cur_irr",
                            sim = 1,
                            dt = "day")
C_lake <- rbind(C_lake1, C_lake2)

C_lake$scenario <- factor(C_lake$scenario, 
                          levels = c("cal", "cur_irr", "no_irr"))
C_lake$lake <- factor(C_lake$lake, levels = lakes)
C_lake      <- C_lake %>%
               mutate_at("date", as.character) %>%
               mutate_at(c("vol_m3", "P_m3", "M_P", "GWin_m3", "M_GWin", "GWout_m3", 
                           "M_GWout", "M_lake"), round)
C_lake$C_lake  <- round(C_lake$C_lake, 2)

# Plot -------------------------------------------------------------------------
ggplot() +
  geom_line(data = C_lake,
            aes(x = as_datetime(.data$date),
                y = .data$C_lake,
                color = .data$scenario)) +
  scale_x_datetime(breaks = seq(as_datetime("1986-01-01"),
                                as_datetime("2021-01-01"),
                                "5 years"),
                   minor_breaks = "1 year",
                   labels = as.character(seq(0, 35, 5))) +
  scale_y_continuous(limits = c(0,40),
                     expand = c(0,0)) +
  scale_color_manual(name = "Scenario",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  facet_wrap(~lake, ncol = 1) +
  labs(x = "Time (yr)", y = "Magnesium (mg/L)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")
```

<br>

### Lake Mg summary statistics

Here, we use the same definitions of "infrequent high" and "infrequent low" as
we do for lake levels, where "infrequent high" equates to the 10% exceedance
probability and "infrequent low" equates to the 90% exceedance probability.

<br>

We are tentatively defining "significant impact" as occuring if:

1. The median Mg concentration under *current irrigation* falls outside the
range of the infrequent high to infrequent low Mg concentration under *no
irrigation*. Under this rule, none of the lakes are significantly impacted based
on these MODFLOW results.
2. The maximum Mg concentration more than doubles under *current irrigation* as
compared to under *no irrigation*. Under this rule, Long Lake is significantly
impacted.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
# Summary metrics --------------------------------------------------------------
C_lake_stats <- C_lake %>%
          group_by(.data$lake, .data$scenario) %>%
          summarise(median = median(.data$C_lake),
                    q10 = quantile(.data$C_lake,
                                          probs = (1 - 0.1),
                                          type = 6,
                                          na.rm = TRUE),
                    q90 = quantile(.data$C_lake,
                                          probs = (1 - 0.9),
                                          type = 6,
                                          na.rm = TRUE),
                    min = round(min(.data$C_lake),2),
                    max = round(max(.data$C_lake),2),
                    .groups = "drop") %>%
          mutate(q10 = round(as.numeric(.data$q10),2),
                 q90 = round(as.numeric(.data$q90),2),
                 range_10_90 = .data$q10 - .data$q90) %>%
          select(.data$lake, .data$scenario, .data$min, .data$q90, .data$median, .data$q10, .data$max, .data$range_10_90)

datatable(C_lake_stats,
          colnames = c("Lake", "Scenario", "Minimum (mg/L)", "Infrequent Low (mg/L)", 
                       "Median (mg/L)", "Infrequent High (mg/L)",
                       "Maximum (mg/L)", "Range (10-90%) (mg/L)"))

```

<br>


## 3. Evaluate robustness of dynamics

To evaluate how sensitive these conclusions are to the volume of groundwater
inflow and outflow estimated by MODFLOW, we test using 50% smaller and 50%
larger groundwater inflow values. We adjust groundwater outflow to keep net
groundwater the same.

<br>

We cannot be sure that the median _measured_ upgradient groundwater
concentration matches the median upgradient groundwater concentration _actually
entering each lake_. But to generate groundwater Mg concentrations closest to
median measured upgradient Mg concentrations, you would need to use:

* Pleasant Lake: 50% larger groundwater inflow
* Long Lake: 50% smaller groundwater inflow
* Plainfield Lake: 50% larger groundwater inflow

<br>

Big-picture solute dynamics are not meaningfully changed by scaling groundwater
flux magnitudes up or down, nor are the conclusions from our quantitative
ecological rules (yay!). I believe we can conclude that given a) the lake level
time series calculated by MODFLOW and b) the assumptions about starting and
ending lake concentrations in WY2018, our approach is robust to uncertainties in
flux magnitude or upgradient groundwater concentrations.

<br>

### Resulting groundwater concentrations

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
MODFLOW_smaller <- CSLSdata::MODFLOW %>%
                   mutate(GWout_m3 = .data$GWout_m3 - abs(.data$GWin_m3/2),
                          GWin_m3 = .data$GWin_m3/2)

MODFLOW_larger <- CSLSdata::MODFLOW %>%
                   mutate(GWout_m3 = .data$GWout_m3 + abs(.data$GWin_m3/2),
                          GWin_m3 = 1.5*.data$GWin_m3)

C_pcpn  <- get_C_pcpn(param, lakes = lakes)
C_GWout <- get_C_lake(param,
                      summary_type = "median",
                      value_name = "C_GWout")

C1_lake <- get_C_lake(param, 
                      start_date = as_datetime("2018-09-01"),
                      end_date = as_datetime("2018-10-31"),
                      summary_type = "median",
                      value_name = "C1_lake")
C2_lake <- get_C_lake(param, 
                      start_date = as_datetime("2019-09-01"),
                      end_date = as_datetime("2019-10-31"),
                      summary_type = "median",
                      value_name = "C2_lake")
C_vals  <- left_join(C_pcpn, C1_lake, by = "lake") %>%
           left_join(C2_lake, by = "lake") %>%
           left_join(C_GWout, by = "lake")


C_GWin_smaller <- calculate_C_GWin(C_vals, fluxes = MODFLOW_smaller) %>%
                  mutate(type = "50% Smaller GW Flux")
C_GWin_larger  <- calculate_C_GWin(C_vals, fluxes = MODFLOW_larger) %>%
                  mutate(type = "50% Larger GW Flux")
C_GWin$type    <- "MODFLOW GW Flux"
C_GWin_new <- rbind(C_GWin, C_GWin_smaller) %>%
              rbind(C_GWin_larger)

datatable(C_GWin_new)
```

<br>

### Dynamics with 50% smaller groundwater fluxes
**Important Note** the y-axis is limited to Mg = 50 mg/L, but values at Long Lake can peak to ~250 mg/L.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
# Smaller 
C_vals <- C_GWin_smaller %>%
          select(lake = .data$lake,
                 C_pcpn = .data$C_pcpn,
                 C0_lake = .data$C_GWout,
                 C_GWin = .data$C_GWin)
C_lake1 <- calculate_C_lake(C_vals,
                            scenario = "no_irr",
                            sim = 1,
                            dt = "day",
                            fluxes = MODFLOW_smaller)
C_lake2 <- calculate_C_lake(C_vals,
                            scenario = "cur_irr",
                            sim = 1,
                            dt = "day",
                            fluxes = MODFLOW_smaller)
C_lake <- rbind(C_lake1, C_lake2)

C_lake$scenario <- factor(C_lake$scenario, 
                          levels = c("cal", "cur_irr", "no_irr"))
C_lake$lake <- factor(C_lake$lake, levels = lakes)
C_lake      <- C_lake %>%
               mutate_at("date", as.character) %>%
               mutate_at(c("vol_m3", "P_m3", "M_P", "GWin_m3", "M_GWin", "GWout_m3", 
                           "M_GWout", "M_lake"), round)
C_lake$C_lake  <- round(C_lake$C_lake, 2)

# Plot -------------------------------------------------------------------------
ggplot() +
  geom_line(data = C_lake,
            aes(x = as_datetime(.data$date),
                y = .data$C_lake,
                color = .data$scenario)) +
  scale_x_datetime(breaks = seq(as_datetime("1986-01-01"),
                                as_datetime("2021-01-01"),
                                "5 years"),
                   minor_breaks = "1 year",
                   labels = as.character(seq(0, 35, 5))) +
  scale_y_continuous(limits = c(0,50),
                     expand = c(0,0)) +
  scale_color_manual(name = "Scenario",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  facet_wrap(~lake, ncol = 1) +
  labs(x = "", y = "Magnesium (mg/L)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")

C_lake_sm <- C_lake %>%
          group_by(.data$lake, .data$scenario) %>%
          summarise(median = median(.data$C_lake),
                    q10 = quantile(.data$C_lake,
                                          probs = (1 - 0.1),
                                          type = 6,
                                          na.rm = TRUE),
                    q90 = quantile(.data$C_lake,
                                          probs = (1 - 0.9),
                                          type = 6,
                                          na.rm = TRUE),
                    min = round(min(.data$C_lake),2),
                    max = round(max(.data$C_lake),2),
                    .groups = "drop") %>%
          mutate(q10 = round(as.numeric(.data$q10),2),
                 q90 = round(as.numeric(.data$q90),2),
                 range_10_90 = .data$q10 - .data$q90) %>%
          select(.data$lake, .data$scenario, .data$min, .data$q90, .data$median, .data$q10, .data$max, .data$range_10_90)

```
<br>


### Dynamics with 50% larger groundwater fluxes
**Important Note** the y-axis is limited to Mg = 40 mg/L, but values at Long Lake can peak to ~135 mg/L.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
# Larger 
C_vals <- C_GWin_larger %>%
          select(lake = .data$lake,
                 C_pcpn = .data$C_pcpn,
                 C0_lake = .data$C_GWout,
                 C_GWin = .data$C_GWin)
C_lake1 <- calculate_C_lake(C_vals,
                            scenario = "no_irr",
                            sim = 1,
                            dt = "day",
                            fluxes = MODFLOW_larger)
C_lake2 <- calculate_C_lake(C_vals,
                            scenario = "cur_irr",
                            sim = 1,
                            dt = "day",
                            fluxes = MODFLOW_larger)
C_lake <- rbind(C_lake1, C_lake2)

C_lake$scenario <- factor(C_lake$scenario, 
                          levels = c("cal", "cur_irr", "no_irr"))
C_lake$lake <- factor(C_lake$lake, levels = lakes)
C_lake      <- C_lake %>%
               mutate_at("date", as.character) %>%
               mutate_at(c("vol_m3", "P_m3", "M_P", "GWin_m3", "M_GWin", "GWout_m3", 
                           "M_GWout", "M_lake"), round)
C_lake$C_lake  <- round(C_lake$C_lake, 2)

# Plot -------------------------------------------------------------------------
ggplot() +
  geom_line(data = C_lake,
            aes(x = as_datetime(.data$date),
                y = .data$C_lake,
                color = .data$scenario)) +
  scale_x_datetime(breaks = seq(as_datetime("1986-01-01"),
                                as_datetime("2021-01-01"),
                                "5 years"),
                   minor_breaks = "1 year",
                   labels = as.character(seq(0, 35, 5)))  +
  scale_y_continuous(limits = c(0,40),
                     expand = c(0,0)) +
  scale_color_manual(name = "Scenario",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  facet_wrap(~lake, ncol = 1) +
  labs(x = "", y = "Magnesium (mg/L)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")

C_lake_lg <- C_lake %>%
          group_by(.data$lake, .data$scenario) %>%
          summarise(median = median(.data$C_lake),
                    q10 = quantile(.data$C_lake,
                                          probs = (1 - 0.1),
                                          type = 6,
                                          na.rm = TRUE),
                    q90 = quantile(.data$C_lake,
                                          probs = (1 - 0.9),
                                          type = 6,
                                          na.rm = TRUE),
                    min = round(min(.data$C_lake),2),
                    max = round(max(.data$C_lake),2),
                    .groups = "drop") %>%
          mutate(q10 = round(as.numeric(.data$q10),2),
                 q90 = round(as.numeric(.data$q90),2),
                 range_10_90 = .data$q10 - .data$q90) %>%
          select(.data$lake, .data$scenario, .data$min, .data$q90, .data$median, .data$q10, .data$max, .data$range_10_90)

```

<br>

### Comparison summary statistics

Does the median Mg concentration under *current irrigation* falls outside the
range of the infrequent high to infrequent low Mg concentration under *no
irrigation*? **No.** Under this rule, none of the lakes are significantly
impacted under normal, 50% higher, or 50% lower MODFLOW groundwater fluxes.

<br>

Does the maximum Mg concentration more than double under *current irrigation* as
compared to under *no irrigation*? **Yes, at Long Lake.** Under this rule,
Pleasant and Plainfield are not significantly impacted under normal, 50% higher,
or 50% lower MODFLOW groundwater fluxes, but Long Lake is under all three
scenarios.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6}
C_lake_sm$type    <- "50% Smaller GW Flux"
C_lake_lg$type    <- "50% Larger GW Flux"
C_lake_stats$type <- "MODFLOW GW Flux"
C_lake_all_stats <- rbind(C_lake_stats, C_lake_sm) %>%
                    rbind(C_lake_lg)

C_lake_all_stats$median      <- round(C_lake_all_stats$median, 2)
C_lake_all_stats$range_10_90 <- round(C_lake_all_stats$range_10_90, 2)

datatable(C_lake_all_stats,
          colnames = c("Lake", "Scenario", "Minimum (mg/L)", "Infrequent Low (mg/L)", 
                       "Median (mg/L)", "Infrequent High (mg/L)",
                       "Maximum (mg/L)", "Range (10-90%) (mg/L)",
                       "Type"))

```

<br>

## 4. Identify thresholds

This is a bonus section, created to try to explore some of the reasons for the
observations above. Long Lake evapoconcentration is clearly tied to lake level
(it spikes below a clearly defined level). The drivers behind Plainfield
dynamics are not immediately obvious.

<br>

### Lake level vs. Lake Mg

**Important Note** the x-axis is limited to Mg = 40 mg/L, but values at Long Lake can peak to ~165 mg/L.

<br>

**Threshold elevation noted here:** the elevation at which lake concentration
begins to exceed the upgradient groundwater concentration. This only occurs at
Long (not Pleasant or Plainfield).

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4}
C_vals <- C_GWin %>%
          select(lake = .data$lake,
                 C_pcpn = .data$C_pcpn,
                 C0_lake = .data$C_GWout,
                 C_GWin = .data$C_GWin)
C_lake1 <- calculate_C_lake(C_vals,
                            scenario = "no_irr",
                            sim = 1,
                            dt = "day")
C_lake2 <- calculate_C_lake(C_vals,
                            scenario = "cur_irr",
                            sim = 1,
                            dt = "day")

C_lake1$lake <- factor(C_lake1$lake, levels = lakes)
C_lake2$lake <- factor(C_lake2$lake, levels = lakes)

yintercept <- C_lake2 %>% 
              filter(.data$C_lake >= .data$C_GWin) %>%
              group_by(.data$lake) %>%
              summarise(level_m = max(.data$level_m)) %>%
              select(.data$lake, .data$level_m) %>%
              unique()
yintercept$lake <- factor(yintercept$lake, levels = lakes)

ggplot() +
  geom_point(data = C_lake2,
             aes(x = .data$C_lake,
                 y = NISTmeterTOft(.data$level_m),
                 color = "cur_irr",
                 shape = "cur_irr"),
             size = 1) +
  geom_point(data = C_lake1,
             aes(x = .data$C_lake,
                 y = NISTmeterTOft(.data$level_m),
                 color = "no_irr",
                 shape = "no_irr"),
             size = 1) +
  geom_hline(data = yintercept, 
             aes(yintercept = NISTmeterTOft(.data$level_m)),
             linetype = "dashed") +
  geom_text(data = yintercept, 
             aes(x = 30,
                 y = NISTmeterTOft(.data$level_m) + 0.5,
                 label = sprintf("%.1f ft", NISTmeterTOft(.data$level_m))),
            family = "Segoe UI Semilight",
            size = 3) +
  scale_x_continuous(limits = c(0,40)) +
  scale_color_manual(name = "Scenario",
                     values = c("steelblue4", "goldenrod1"),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  scale_shape_manual(name = "Scenario",
                     values = c(16, 17),
                     breaks = c("no_irr", "cur_irr"),
                     labels = c("No Irrigation",
                                "Current Irrigation")) +
  facet_wrap(~lake, ncol = 3, scales = "free") +
  labs(x = "Magnesium (mg/L)", y = "Lake Level (ft)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")

```

<br>

### Investigate Plainfield more closely

Plainfield Lake patterns are a little more difficult to uncover. Here we try
looking at the difference between "No Irrigation" and "Current Irrigation" Mg
values plotted vs. both "No Irrigation" and "Current Irrigation" lake levels to
see if it helps unearth patterns. Patterns remain elusive to me.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4}
C_lake <- left_join(C_lake1, C_lake2, 
                    by = c("lake", "date")) %>%
          mutate(diff = .data$C_lake.y - .data$C_lake.x,
                 water_year = ifelse(month(.data$date) >= 10,
                                     year(.data$date) + 1,
                                     year(.data$date))) %>%
          select(.data$lake, .data$date, .data$level_m.x, 
                 .data$level_m.y, .data$water_year,
                 .data$C_lake.x, .data$C_lake.y, .data$diff,
                 .data$vol_m3.x, .data$vol_m3.y)

ggplot() +
  geom_point(data = C_lake,
             aes(x = .data$diff,
                 y = NISTmeterTOft(.data$level_m.x)),
             size = 1) +
  facet_wrap(~lake, ncol = 3, scales = "free") +
  labs(x = "Magnesium Difference (mg/L)", y = "No Irrigation Lake Level (ft)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"))

```

<br>


```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4}
ggplot() +
  geom_point(data = C_lake,
             aes(x = .data$diff,
                 y = NISTmeterTOft(.data$level_m.y)),
             size = 1) +
  facet_wrap(~lake, ncol = 3, scales = "free") +
  labs(x = "Magnesium Difference (mg/L)", y = "Current Irrigation Lake Level (ft)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"))
```

<br>

Just in case something is clearer by looking at the timeseries, here's
Plainfield lake levels and differences. No surprise, still kind of messy. You're
more likely to see evapoconcentrating effects at the lower lake levels (positive
Mg difference aka current irrigation Mg is higher), but there are unexplained dips
to loss of source (negative Mg difference aka current irrigation Mg is lower)
throughout that period as well.

<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=4}
C_lake$level_ft.y <- NISTmeterTOft(C_lake$level_m.y)
C_lake$level_ft.x <- NISTmeterTOft(C_lake$level_m.x)
ggplot(filter(C_lake, .data$lake == "Plainfield")) +
  geom_line(aes(x = .data$date,
                 y = .data$diff,
                color = "Mg Difference"),
            size = 1) +
  geom_line(aes(x = .data$date,
                 y = (.data$level_ft.y-1096),
                color = "Cur. Irr. Lake Level")) +
  geom_line(aes(x = .data$date,
                 y = (.data$level_ft.x-1096),
                color = "No Irr. Lake Level")) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # facet_wrap(~lake, ncol = 1, scales = "free") +
  scale_y_continuous(sec.axis = sec_axis(~.+1096,
                     name = "Current Irrigation Lake Level (ft)")) +
  labs(x = "", y = "Magnesium Difference (mg/L)", color = "") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")
```

<br>
<br>

Is there instead a relationship between the Mg difference and Precipitation
minus Evaporation at Plainfield? Also difficult to say, no matter whether you
break it down by water year or by season.

<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.5}
MODFLOW_PE <- CSLSdata::MODFLOW %>%
              filter(!is.na(.data$GWin_m3),
                     .data$scenario != "cal",
                     .data$sim == 1) %>%
              mutate(month = month(.data$date),
                     year = year(.data$date),
                     water_year = ifelse(.data$month >= 10,
                                         .data$year + 1,
                                         .data$year),
                     water_year_date = as_datetime(sprintf("%d-04-01",
                                                           .data$water_year)),
                     PE_m3 = abs(.data$P_m3)-abs(.data$E_m3)) %>%
              group_by(lake = .data$lake, 
                       date = .data$water_year_date) %>%
              summarise(PE_m3 = sum(.data$PE_m3),
                        .groups = "drop")

ggplot() +
  geom_col(data = filter(MODFLOW_PE, .data$lake == "Plainfield"),
            aes(x = .data$date,
                y = NISTcubMeterTOacreFt(.data$PE_m3)/50),
           color = "skyblue",
           fill = "skyblue",
           alpha = 0.4)  +
  geom_line(data = filter(C_lake, .data$lake == "Plainfield"),
             aes(x = .data$date,
                 y = .data$diff)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_y_continuous(sec.axis = sec_axis(~.*50, name = "P - E (ac-ft)")) +
  # facet_wrap(~lake, ncol = 1, scales = "free") +
  labs(x = "", y = "Magnesium Difference (mg/L)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")


```
<br>

```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=3.5}
MODFLOW_PE <- CSLSdata::MODFLOW %>%
              filter(!is.na(.data$GWin_m3),
                     .data$scenario != "cal",
                     .data$sim == 1) %>%
              mutate(month = month(.data$date),
                     year = year(.data$date),
                     water_year = ifelse(.data$month >= 10,
                                         .data$year + 1,
                                         .data$year),
                     season_year = ifelse(.data$month == 12,
                                          .data$year + 1,
                                          .data$year),
                     season_month = ifelse(.data$month %in% c(12,1,2),
                                     1, ifelse(.data$month %in% c(3,4,5),
                                                      4, ifelse(.data$month %in% c(6,7,8),
                                                                       7, 10))),
                     season_date = as_datetime(sprintf("%d-%d-01", .data$season_year, 
                                           .data$season_month)),
                     PE_m3 = abs(.data$P_m3)-abs(.data$E_m3)) %>%
              group_by(lake = .data$lake, 
                       date = .data$season_date,
                       season_month = .data$season_month) %>%
              summarise(PE_m3 = sum(.data$PE_m3),
                        .groups = "drop")

ggplot() +
  geom_col(data = filter(MODFLOW_PE, .data$lake == "Plainfield"),
            aes(x = .data$date,
                y = NISTcubMeterTOacreFt(.data$PE_m3)/10,
                fill = as.factor(.data$season_month)),
           color = NA,
           alpha = 0.5)  +
  geom_line(data = filter(C_lake, .data$lake == "Plainfield"),
             aes(x = .data$date,
                 y = .data$diff)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_y_continuous(sec.axis = sec_axis(~.*10, name = "P - E (ac-ft)")) +
  # scale_color_brewer(name = "Season",
  #                    labels = c("Winter", "Spring", "Summer", "Fall"),
  #                    palette = "Dark2") +
  scale_fill_brewer(name = "Season",
                     labels = c("Winter", "Spring", "Summer", "Fall"),
                     palette = "Dark2") +
  # facet_wrap(~lake, ncol = 1, scales = "free") +
  labs(x = "", y = "Magnesium Difference (mg/L)") +
  theme_bw() +
  theme(text = element_text(family = "Segoe UI Semilight"),
        legend.position = "top")


```
